Cannibalism among the supermassive black holes

### 2019-02-18 paper.tex
More edits

\begin{abstract}
We perform N-body simulations on 13 galaxies extracted from a cosmological simulation of hierarchical structure formation with total mass $10^{12} \msun < M < 10^{13} \msun$ from $4 \ge z \ge 0$.  We track the dynamical evolution of the orbiting black holes (BHs) around their host central BH.  Of the 120 orbiting BHs, we find that 50 merge with their host central BHs, 11 are ejected from their respective host galaxies, and 59 are still orbiting at the end of the simulations.  48 BHs are kicked to a higher orbit after close interaction with the central BH, after which none of them merge with their host.  We calculate the x-ray luminosities of the orbiting BHs at z=0, and find a range of $10^{18}-10^{37}$ $erg\ s^{-1}$.
\end{abstract}

@@

How common are mergers of massive black holes?  For the galaxies themselves we know that minor mergers are frequent from the observed (\citet{2010ApJ...718L..73V}, \citet{2008ApJ...677L...5V}) evolution of the size and mass of these systems.  But assuming that all galaxy mergers lead to SMBH mergers overpredicts the observations of gravitational waves from the pulsar timing arrays (PTAs) (\citet{2008MNRAS.390..192S}, \citet{2009MNRAS.394.2255S}, \citet{2013MNRAS.433L...1S}, \citet{2014ApJ...789..156M}, \citet{2015ApJ...799..178K}, \citet{2018ApJ...856...42S}, \citet{2018arXiv180403143I}).  But we do not often see multiple SMBHs at the centers of massive ellipticals.  So, what does happen?  One paper (\citet{2018MNRAS.473.3410R}) has indicated that dynamical interactions among multiple black holes, which eject a non-negligible fraction of the mass, may solve this problem.  The present paper also addresses this purported dynamical solution, focusing attention on the large fraction of lower mass black holes that remain to be detected as they orbit in massive galaxies.  What we argue in this paper is that as a result of the N-body interactions among the infalling black holes some are ejected, some remain in extended orbits and some (the few most massive ones) do in fact merge, but these do not exceed the PTA limits, and the orbiting ones should be detectable via their ability to accrete gas and emit radiation.

@@

The net black hole mass accretion rate is approximated by [DO I REFERENCE INAYOSHI, OR ANOTHER SOURCE?]:
\begin{equation}
    \dot{M}_{acc} = \frac{4{\pi}G^2M^2\rho}{(c_s^2+v^2)^{3/2}} \frac{c_s^2+v^2}{c^2} \text{  M}_\odot \text{ Myr}^{-1},
\end{equation}
where M is black hole mass, G is the gravitational constant, and v is the black hole velocity.

@@

The Eddington limit is known to be:
\begin{equation}
    L_{edd} = \num{1.26e38} (M/M_{\odot}) \text{  erg s}^{-1}.
\end{equation}

@@

\begin{figure}[htp]
\begin{center}
\includegraphics[width=1.0\columnwidth]{plots/Galaxy_Gas_Compared_to_Babyk.png}
\caption{X-ray luminosity of gas in our galaxies, compared to those from \citet{2018ApJ...857...32B}.}
\label{fig:galgas1}
\end{center}
\end{figure}

>>>>

\begin{abstract}
We perform N-body simulations on 13 galaxies extracted from a cosmological simulation of hierarchical structure formation with total mass $10^{12} \msun < M < 10^{13} \msun$ from $4 \ge z \ge 0$.  We track the dynamical evolution of the orbiting black holes (BHs) around their host central BH.  Of the 120 orbiting BHs, we find that 50 merge with their host central BHs, 11 are ejected from their respective host galaxies, 59 are still orbiting at the end of the simulations.  Across all galaxies, 48 BHs are kicked to a higher orbit after close interaction with the central BH, after which none of them merge with their hosts.  We calculate the x-ray luminosities of the orbiting BHs at z=0, and find a range of $10^{18}-10^{37}$ $erg\ s^{-1}$.
\end{abstract}

@@

How common are mergers of massive black holes?  For the galaxies themselves we know that minor mergers are frequent from the observed (\citet{2010ApJ...718L..73V}, \citet{2008ApJ...677L...5V}, \citet{2019MNRAS.484..595M}) evolution of the size and mass of these systems, but assuming that all galaxy mergers lead to SMBH mergers overpredicts the observations of gravitational waves from the pulsar timing arrays (PTAs) (\citet{2008MNRAS.390..192S}, \citet{2009MNRAS.394.2255S}, \citet{2013MNRAS.433L...1S}, \citet{2014ApJ...789..156M}, \citet{2015ApJ...799..178K}, \citet{2018ApJ...856...42S}, \citet{2018arXiv180403143I}).  However, we do not often see multiple SMBHs at the centers of massive ellipticals.  So, what does happen?  One paper (\citet{2018MNRAS.473.3410R}) has indicated that dynamical interactions among multiple black holes, which eject a non-negligible fraction of the mass, may solve this problem.  The present paper also addresses this purported dynamical solution, focusing attention on the large fraction of lower mass black holes that remain to be detected as they orbit in massive galaxies.  What we argue in this paper is that, as a result of the N-body interactions among the infalling black holes, some are ejected, some remain in extended orbits and some (the few most massive ones) do in fact merge, but these do not exceed the PTA limits, and the orbiting ones should be detectable via their ability to accrete gas and emit radiation.

@@

The net black hole mass accretion rate is approximated by \citet{2018MNRAS.476.1412I} and \citet{2019arXiv190202349I}:
\begin{equation}
    \dot{M}_{acc} = \frac{4{\pi}G^2M^2\rho}{(c_s^2+v^2)^{3/2}} \frac{c_s^2+v^2}{c^2} \text{  M}_\odot \text{ Myr}^{-1},
\end{equation}
where M is black hole mass, G is the gravitational constant, and v is the black hole velocity.

@@

The Eddington limit is defined as:
\begin{equation}
    L_{edd} = \num{1.26e38} (M/M_{\odot}) \text{  erg s}^{-1}.
\end{equation}

@@

\begin{figure}[htp]
\begin{center}
\includegraphics[width=1.0\columnwidth]{plots/Lx_our_Galaxies_compared_to_Babyk_data.png}
\caption{X-ray luminosity of gas in our galaxies, compared to those from \citet{2018ApJ...857...32B}.}
\label{fig:galgas1}
\end{center}
\end{figure}
